Refunctionalization of Abstract Abstract Machines (Functional Pearl)

### 2018-06-23 paper/paper.tex
abstract revised

  In this paper, we show their functional correspondence and how to syntactically 
  transform small-step abstract abstract machines into big-step abstract  definitional
  interpreters.
  The transformations include linearization, fusing, disentangling, refunctionalizing,
  and un-CPS to direct-style with delimited control. Linearization expresses
  non-deterministic choices by first-order data types, after which refunctionalization
  sequentializes the evaluation order by higher-order functions.
  All transformations properly handle the collecting semantics and the nondeterminism 
  of abstract interpretation.

  Following the idea that in deterministic languages, evaluation contexts of reduction
  semantics are defunctionalized continuations, we further show that in nondeterministic 
  languages, evaluation contexts are refunctionalized to extended continuations style.
  Remarkably, we reveal how precise call/return matching in control-flow analysis is 
  obtained by refunctionalizing a small-step abstract abstract machine with proper caching.
\end{abstract}

@@

In fact, what we started is an AAM with an unbounded stack which already precisely 
matches the calls and returns.
Refunctionalization makes the interpreter be a ``big-step'' semantics artifact, 
then the consequence is we have no places to save the context information. 
But refunctionalization not only forces us to sequentialize the order non-deterministic 
evaluation through higher-order functions representing continuations, 
but also drives us to use a new caching algorithm.

@@

\textit{Defunctionalization and Refunctionalization.}
Defunctionalization and refunctionalization build connections between abstract
machines and interpreters.
\citeauthor{Reynolds:72} first shows defunctionalization as a program transformation 
technique that can be used to transform higher-order functions to first-order functions
when specifying a programming language by an interpreter \cite{Reynolds:72, Reynolds:HOSC98-revisited}.
Refunctionalization is the reverse of this transformation.

>>>>

  In this paper, we show their functional correspondence and how to syntactically 
  transform small-step abstract abstract machines into big-step abstract definitional
  interpreters.
  The transformations include linearization, lightweight fusion, disentanglement, 
  refunctionalization, and finally to direct-style with delimited control. 
  Linearization expresses nondeterministic choices by first-order data types, 
  after which refunctionalization transforms the datatype representing continuations
  into higher-order functions.
  Following the idea that in deterministic languages, reduction contexts of reduction
  semantics are defunctionalized continuations, we further apply this idea to AAM --
  reduction contexts of it are refunctionalized to extended continuation-passing style.

  All transformations properly handle the collecting semantics and the nondeterminism 
  of abstract interpretation. Remarkably, we reveal how precise call/return matching in 
  control-flow analysis can be obtained by refunctionalizing a small-step abstract abstract 
  machine with proper caching.
\end{abstract}

@@

In fact, what we started is an AAM with an unbounded stack which already precisely 
matches the calls and returns.
Refunctionalization makes the interpreter be a ``big-step'' semantics artifact, 
then the consequence is we have no places to save the context information. 
But refunctionalization not only forces us to sequentialize the order nondeterministic 
evaluation through higher-order functions representing continuations, 
but also drives us to use a new caching algorithm.

@@

\textit{Defunctionalization and Refunctionalization.}
Defunctionalization and refunctionalization build connections between abstract
machines and interpreters.
\citeauthor{Reynolds:72} first shows defunctionalization as a program transformation 
technique that can be used to transform higher-order functions to first-order functions
when specifying a programming language by an interpreter \cite{Reynolds:72, Reynolds:HOSC98-revisited}.
Refunctionalization is the left inverse of this transformation.
